---
title: "Lab 1: Introduction to Data Analysis in R"
subtitle: "Historical birth records, data transformation, and visualization"
execute:
  eval: false
---

[Download the bare-bones student report](../../templates/lab01-template.qmd){.btn .btn-primary download="lab01-template.qmd"}

## The question

Do birth records show a stable difference between the numbers of boys and girls born each year? We will use a historical dataset compiled by John Arbuthnot and then examine more recent U.S. records.

The purpose is not merely to reproduce calculations. It is to practice a complete analysis cycle:

1. ask a clear question;
2. inspect the data;
3. transform variables when necessary;
4. visualize the relevant pattern;
5. calculate a useful summary; and
6. explain what the evidence does—and does not—show.

## Setup

```{r}
library(dplyr)
library(ggplot2)

load(url("https://raw.githubusercontent.com/GarciaRios/lbj_ug_quant/main/data/arbuthnot.RData"))
```

The `load()` command creates a data frame named `arbuthnot`.

## 1. Inspect the data

```{r}
arbuthnot
dim(arbuthnot)
names(arbuthnot)
glimpse(arbuthnot)
```

Each row is a year. The columns contain the year and the recorded numbers of boys and girls baptized in London.

Answer before moving on:

- What is the unit of observation?
- How many observations and variables are present?
- What years are covered?
- What limitations might baptism records have as a measure of births?

Individual variables can be accessed with `$`:

```{r}
arbuthnot$boys
arbuthnot$girls
```

## 2. Visualize change over time

A line graph is appropriate because year has a meaningful order.

```{r}
ggplot(arbuthnot, aes(x = year, y = girls)) +
  geom_line(color = "#BF5700", linewidth = 0.8) +
  labs(
    x = NULL,
    y = "Recorded baptisms",
    title = "Girls recorded in London baptism data"
  ) +
  theme_minimal()
```

Describe the overall pattern. Look for direction, unusual years, and changes in variability. Avoid explaining causes that the data cannot establish.

Now construct the corresponding graph for boys.

## 3. Calculate totals and create variables

The total number of girls in the records is:

```{r}
sum(arbuthnot$girls)
```

To compare births within each year, create total births and the proportion recorded as boys:

```{r}
arbuthnot <- arbuthnot %>%
  mutate(
    total = boys + girls,
    boy_prop = boys / total
  )
```

Inspect the new variables:

```{r}
select(arbuthnot, year, total, boy_prop)
```

Why is a proportion often easier to compare across years than a raw difference?

## 4. Evaluate a claim

Arbuthnot argued that more boys than girls were born in every year of his records. We can translate that statement into a logical comparison.

```{r}
arbuthnot <- arbuthnot %>%
  mutate(more_boys = boys > girls)

count(arbuthnot, more_boys)
```

A logical variable is either `TRUE` or `FALSE`. The code checks the claim year by year; it does not establish why the pattern occurs.

Visualize the proportion of boys:

```{r}
ggplot(arbuthnot, aes(x = year, y = boy_prop)) +
  geom_hline(yintercept = 0.5, color = "gray60", linetype = "dashed") +
  geom_line(color = "#172A46", linewidth = 0.8) +
  labs(
    x = NULL,
    y = "Proportion recorded as boys",
    title = "Share of recorded births classified as boys"
  ) +
  theme_minimal()
```

Explain the dashed line and summarize the substantive pattern in two or three sentences.

## On Your Own: More recent U.S. births

Load the second dataset:

```{r}
load(url("https://raw.githubusercontent.com/GarciaRios/lbj_ug_quant/main/data/present.RData"))
glimpse(present)
```

Complete the following in the downloadable report:

1. Identify the unit of observation, years covered, and variables.
2. Create `total` and `boy_prop`.
3. Make a clear line graph of total births over time.
4. Make a line graph of `boy_prop` with a reference line at 0.5.
5. Determine whether boys outnumber girls in every observed year.
6. Write a short interpretation comparing the recent pattern with the Arbuthnot records.

A strong interpretation states the pattern, uses relevant quantities, and acknowledges what these descriptive data cannot explain.
